---
title: "Bonanza Creek Hares Report"
subtitle: "Assignment 4 - Task 2"
author: "Karina Johnston"
date: "11/3/2021"
output: 
  html_document: 
    toc: yes
    theme: cerulean
---

```{r setup, include = FALSE, warning = FALSE, message = FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)

# Attach packages
library(tidyverse)
library(here)
library(equatiomatic)
library(lubridate)
library(kableExtra)
```

## Introduction

Bonanza Creek Long Term Ecological Research (LTER) program is part of the US LTER Network. In this report, the counts and sizes of juvenile snowshoe hares recorded from 1998 - 2012 will be explored. Snowshoe hares, *Lepus americanus*, are a 'keystone' prey species in northern boreal forests (Kielland et al. 2017). **INSERT SENTENCES ABOUT WHAT IS IN THIS REPORT.**


```{r}
# read in the data from the .csv file
bonanza_hares <- read_csv(here("data", "bonanza_hares.csv"))
```


## Data and Analyses

Capture-recapture studies were conducted on snowshoe hares at five sites in the Tanana Valley. 


#### Juvenile hare initial investigation

```{r, fig.cap = "**Figure 1.** Juvenile Bonanza Creek hare counts by year (1998 - 2012). Data: Kielland et al. 2017.", fig.align = "center"}
bonanza_newdates <- bonanza_hares %>%  # creates a new dataframe beginning with bonanza_hares
  mutate(date_new = mdy(date)) %>%  # creates a new column using `lubridate::mdy()`
  mutate(year_new = year(date_new))  # creats a new year column

bonanza_juveniles <- bonanza_newdates %>% # creates a new dataframe filtering for juveniles
  filter(age == "j")

bonanza_stats <- bonanza_juveniles %>% # groups by year and counts the number by year
  group_by(year_new) %>% 
  summarize(count = n())

# this code chunk outputs a visualization bar graph for the counts by year data
bonanza_count <- ggplot(data = bonanza_stats, aes(x = year_new, y = count)) +
  geom_col(fill = "cyan4") + # so the bars are not grey
  theme_classic() + # for a cleaner graph with no gridlines
  labs(x = "Year", y = "Number of juvenile hares") + # to have formal axis labels 
  scale_x_continuous(breaks = 1999:2012) # to clean up the x-axis labels

bonanza_count

max_count <- max(bonanza_stats$count)
min_count <- min(bonanza_stats$count)
mean_count <- mean(bonanza_stats$count)
sd_count <- sd(bonanza_stats$count)
median <- median(bonanza_stats$count)
```

Distribution of juvenile hare counts was positively skewed. 

`r max_count`


Mean annual number of juvenile hares trapped was XXX $\pm$ XXXX (mean $\pm$ SD)

#### Annual juvenile hare trap counts



#### Visualize juvenile hare weights

weight_male <- hares %>% # creates a new dataframe 
filter(sex == "m") %>%# to filter male weights only
drop_na(weight) # to drop the NAs
pull(weight) # to pull a vector of just weights



#### Juvenile weight comparison 



#### Relationship between juvenile weight and hind foot length



## Summary of Findings



## Citations

Kielland, K., F.S. Chapin, R.W. Ruess, and Bonanza Creek LTER. 2017. Snowshoe hare physical data in Bonanza Creek Experimental Forest: 1999-Present ver 22. Environmental Data Initiative. https://doi.org/10.6073/pasta/03dce4856d79b91557d8e6ce2cbcdc14


