---
title: "Assignment 4 - Task 2 - Bonanza Creek LTER Hares"
author: "Karina Johnston"
date: "11/3/2021"
output: 
  html_document: 
    toc: yes
    theme: cerulean
---

```{r setup, include = TRUE, warning = FALSE, message = FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)

# Attach packages
library(tidyverse)
library(here)
library(equatiomatic)
```

## Introduction

Bonanza Creek LTER is part of the US Long Term Ecological Research Network. In this report, the counts and sizes of juvenile snowshoe hares recorded from 1998 - 2012 will be explored. 


```{r}
# read in the data from the .csv file
bonanza_hares <- read_csv(here("data", "bonanza_hares.csv"))
```


## Data and Analyses



### Juvenile Hare Initial Investigation



### Annual juvenile hare trap counts



### Visualize juvenile hare weights



### Juvenile weight comparison 



### Relationship between juvenile weight and hind foot length



## Summary of Findings



## Citations

Kielland, K., F.S. Chapin, R.W. Ruess, and Bonanza Creek LTER. 2017. Snowshoe hare physical data in Bonanza Creek Experimental Forest: 1999-Present ver 22. Environmental Data Initiative. https://doi.org/10.6073/pasta/03dce4856d79b91557d8e6ce2cbcdc14


